Homicide Increases in the COVID-19 Era

### 2025-02-24 Draft1.qmd
Updated draft based on report feedback about moving code blocks to the end

author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
fontsize: 10pt
toc: true                   # add table of contents at the beginning
toc-depth: 3               # Only titles that start with # or ##
geometry: margin=1in
bibliography: references.bib
format:
  pdf: default
---

@@

# Loading in packages and dataset

```{r, output=FALSE}
library(sf) # for map
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer) # for map color scheme
library(tidygeocoder) # to get zip codes of addresses

@@

```{r}

@@

```{r}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

```{r}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

# Plot bar chart with horizontal labels
barplot(crime_counts, col = "lightblue",
        main = "Crime Occurrences by Day of the Week",
        xlab = "Day of the Week", ylab = "Number of Crimes",
        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
```

@@

```{r}
# Remove missing values and empty strings in Vict.Sex column
crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))

@@

```{r}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

```{r}
# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)

@@

### Interpretation
The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased. However, there are some that a typical family is at or below the poverty line. We intend to investigate if there is a spatial correlation between these poorer zip codes and the incidence of crimes, particularly homicide, and whether this is affected by pandemic conditions.

@@

```{r}
# using latitude and longitude to obtain the zip codes of the crime locations
# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))

@@

### Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code. However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions. Future visualizations and analyses could consider pandemic vs non-pandemic years as variables.

@@

## Using Regression to Examine Weekly Crime Rates

```{r}
weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])

@@

summary(ar1_trend_regression)
```
### Interpretation

@@

```{r}
lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)

summary(lm_income_crimes)
```
### Interpretation

Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed. The r^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes. As stated with the scatterplot, a larger sample and considering other factors might yield different results.

# Bibliography

>>>>

author: "Marco Stine, Kevin Hu, Bianca Schutz, and Kaitlyn Kirt"
fontsize: 10pt
toc: true                   # add table of contents at the beginning
toc-depth: 3               # Only titles that start with # or ##
geometry: margin=1in
bibliography: references.bib
format:
  pdf: default
number-sections: true
---

@@

```{r, output=FALSE, echo=FALSE}
library(sf) # for map
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer) # for map color scheme
library(tidygeocoder) # to get zip codes of addresses

@@

```{r, echo=FALSE}

@@

```{r, echo=FALSE}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

```{r, echo=FALSE, fig.cap = "Crime Occurrences by Day of the Week"}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

# Plot bar chart with horizontal labels
barplot(crime_counts, col = "lightblue",
        xlab = "Day of the Week", ylab = "Number of Crimes",
        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
```

@@

```{r, echo=FALSE}
# Remove missing values and empty strings in Vict.Sex column
crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))

@@

```{r, echo=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

```{r, echo=FALSE}
# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)

@@

The maps demonstrate that median income in certain LA zip codes (particularly those to the west) have a much higher median household income than the rest of the city, but over the last six years, the median income in most other LA zip codes has increased. However, there are some that a typical family is at or below the poverty line. We intend to investigate if there is a spatial correlation between these poorer zip codes and the incidence of crimes, particularly homicide, and whether this is affected by pandemic conditions.

@@

```{r, echo=FALSE}
# using latitude and longitude to obtain the zip codes of the crime locations
# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))

@@

Based on this plot, we do not observe a clear trend between the number of crimes committed and the median income in that zip code. However, this scatterplot only accounts for a small portion of the dataset and averages across pre- and post-pandemic conditions. Future visualizations and analyses could consider pandemic vs non-pandemic years as variables.

@@

## Regression to Examine Weekly Crime Rates

```{r, echo=FALSE}
library(broom)

weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])

@@

ar1_trend_regression %>%
tidy() %>%
  mutate(
    p.value = scales::pvalue(p.value)
  ) %>%
  knitr::kable(
    caption = "Linear Regression Model Results Estimating Crimes Per Week",
    col.names = c("Predictor", "B", "SE", "t", "p"),
    digits = c(0, 2, 3, 2, 3),
    align = c("l", "r", "r", "r", "r")
  )
```

@@

```{r, echo=FALSE}
lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)

summary(lm_income_crimes)

lm_income_crimes %>%
tidy() %>%
  mutate(
    p.value = scales::pvalue(p.value)
  ) %>%
  knitr::kable(
    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
    col.names = c("Predictor", "B", "SE", "t", "p"),
    digits = c(0, 2, 3, 2, 3),
    align = c("l", "r", "r", "r", "r")
  )
```

Based on the p-value, we observe that median income is not significant at a 0.05 level, meaning it does not significantly decrease the amount of crimes committed. The r\^2 value additionally tells us that the income only explains a small portion of the variation in the number of crimes. As stated with the scatterplot, a larger sample and considering other factors might yield different results.

# References {.unnumbered}

::: {#refs}
:::

# Appendix {.unnumbered}

```{r, eval=FALSE}
library(sf) # for map
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer) # for map color scheme
library(tidygeocoder) # to get zip codes of addresses

## Figure 1 - Examining Crime Density from 2010–2019 and 2020-Present {.unnumbered}

```{r, eval=FALSE}

## Figure 2 - Examining Number Of Crimes Over Time {.unnumbered}

```{r, eval=FALSE}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

## Figure 3 - Crime rate frequency by day of the week {.unnumbered}

```{r, eval=FALSE}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

# Plot bar chart with horizontal labels
barplot(crime_counts, col = "lightblue",
        xlab = "Day of the Week", ylab = "Number of Crimes",
        names.arg = names(crime_counts), las = 1, cex.names = 0.8, cex.axis = 0.6)
```

## Figure 4 - Gender Distribution for crimes over time {.unnumbered}

```{r, eval=FALSE}
# Remove missing values and empty strings in Vict.Sex column
crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))

## Figure 5 - Median Household Income by Zip Code in Los Angeles by Year {.unnumbered}

```{r, eval=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

```{r, eval=FALSE}
# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)

## Figure 6 - Crime Rates Compared to Median Household Income {.unnumbered}

```{r, eval=FALSE}
# using latitude and longitude to obtain the zip codes of the crime locations
# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))

## Crime \~ Week Regression Code {.unnumbered}

```{r, eval=FALSE}
library(broom)

weekly_counts$CrimesLag <- c(NA, weekly_counts$Crimes[-nrow(weekly_counts)])

ar1_trend_regression %>%
tidy() %>%
  mutate(
    p.value = scales::pvalue(p.value)
  ) %>%
  knitr::kable(
    caption = "Linear Regression Model Results Estimating Crimes Per Week",
    col.names = c("Predictor", "B", "SE", "t", "p"),
    digits = c(0, 2, 3, 2, 3),
    align = c("l", "r", "r", "r", "r")
  )
```

## Crime \~ Income Regression Code {.unnumbered}

```{r, eval=FALSE}
lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)

summary(lm_income_crimes)

lm_income_crimes %>%
tidy() %>%
  mutate(
    p.value = scales::pvalue(p.value)
  ) %>%
  knitr::kable(
    caption = "Linear Regression Model Results Estimating Crimes Based on Median Income",
    col.names = c("Predictor", "B", "SE", "t", "p"),
    digits = c(0, 2, 3, 2, 3),
    align = c("l", "r", "r", "r", "r")
  )
```
